surfing

### 2021-05-07 inst/surfing.Rmd
better writing

Competitive surfing involves surfers competing against one or two
other surfers in heats lasting 30-50 minutes.  Points are awarded on a
10-point scale: each of five judges awards points to a surfer riding a
particular wave.

The five judges rate each of the waves ridden by a surfer and award
points based on a range of subjective criteria including commitment
and degree of difficulty, innovation and progress of manoeuvres,
combination of major manoeuvre, variety of manoeuvres and speed, power
and flow.  The top score and bottom score among the five judges are
removed and the remaining three judges' scores are averaged to give
the final score for the surfer for that wave.

Each surfer's aggregate score is given by the average of the two
highest-scoring waves.  The winner of the tournament is the surfer
with the highest aggregate score.

The World Surf League (WSL) is the main governing body of surfing
competitions.  The WSL conducts a world tour in which the best ranked
34 surfers compete; in addition, at each tour venue, two "wildcard"
surfers also enter the competition.

The 2019 WSL tour saw competitive events held at eleven different surf
locations.  In each of these locations the wave type of wave
differs.  Some events are known to be more favourable to those who have
a high number of manoeuvres at their disposal and others are more
suited to surfers who regularly surf bigger waves.

@@

```{r showsurfingmaxppie, echo=FALSE, fig.cap="dot chart of result"}
dotchart(surfing_maxp)
dotchart(log10(surfing_maxp))
pie(surfing_maxp)
```

>>>>

Surfing is a popular and growing activity in terms of both
participation and viewing.  Competitive surfing involves surfers
competing against one or two other surfers in heats lasting 30-50
minutes.  Points are awarded on a 10-point scale: each of five judges
awards points to a surfer riding a particular wave.

The typical format is as follows.  Up to four surfers---the
competitors---are in the water simultaneously, watched by up to five
judges.  Subject to surfing etiquette, competitors are free to catch a
wave at their discresion within the heat's time window, typically
20-30 minutes.  The judges rate each wave ride and award points to the
surfer based on a range of subjective criteria including commitment
and degree of difficulty, innovation and progress of manoeuvres,
variety of manoeuvres and speed, power and flow.  The top score and
bottom score among the five judges are removed and the remaining three
judges' scores are averaged to give the final score for the surfer for
that wave.  Each surfer's aggregate score is given by the average of
the two highest-scoring waves.  The winner of the tournament is the
surfer with the highest aggregate score.

Such scoring systems are designed to account for the random nature of
wave quality while reflecting competitors' abilities fairly.
Differences in wave quality between successive days mean that direct
comparison of scores between one day and another are not informative
about competitors' abilities as they are strongly dependent on details
of wave quality at the time of the competition.  However, for a
particular heat, the order statistic---that is, which competitor
scored most highly, second highest, and third---is informative about
competitors' abilities: the wave environment is common to each surfer.

The World Surf League (WSL) is the main governing body of surfing
competitions.  The WSL conducts a world tour in which the best ranked
34 surfers compete; in addition, at each tour venue, two "wildcard"
surfers also enter the competition.  The 2019 WSL tour saw competitive
events held at eleven different surf locations.  In each of these
locations the wave type of wave differs.  Some events are known to be
more favourable to those who have a high number of manoeuvres at their
disposal and others are more suited to surfers who regularly surf
bigger waves.

@@

```{r showsurfingmaxppie, echo=FALSE, fig.height=8,fig.cap="dot chart of result"}
dotchart(surfing_maxp)
dotchart(log10(surfing_maxp))
pie(surfing_maxp)
```
